Technologies of Semantic WEB as an environment of application development and integration

### 2021-12-05 paper-aicts-cherkashin.tex
Presentation should be ok now

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Semantic Web technologies and standards are becoming one of productive assets for software development and integration on various levels of design and implementation.  This include OSI level 6 data representation, database and file storage formats, the source data for indexes, hypertext markup, source data for document authoring, representing abstract software models, \emph{etc}.  They also smoothly combined with development and data analysis tools, giving raise a new common-ground environment for realizing integration on the semantic level.

@@

KGs changed the view point to the knowledge base as being a part of whole totality of knowledge, implying the obeying the global standards and techniques of its acquisition and processing \cite{hogan}.  The following outlines their definitions \cite{hogan}.
 \begin{itemize}
  \item Notions \emph{data} and \emph{knowledge} are converged into ``something is \emph{known}'';
  \item KG contains, in general, data, relations, and metadata;
  \item The same KG is being filled in and processed by different independent modules, \emph{e.g.}, withing SPARQL queries with UPDATE statements;
  \item KG fundamental software development allow to \emph{postpone} the formal definition of a data schema;
  \item There is three points of view to graph schemata: \emph{semantic}, which is aimed at generalization, \emph{validating}, reflecting semantic verification, checking completeness w.r.t. sets of predefined general relations, and \emph{emergent}, observed or deduced from exploitation experience, as a set of new generalized structures and \emph{refactoring} of the KG.
  \end{itemize}

\begin{figure}[bth]
  \begin{center}
    \includegraphics[width=0.9\linewidth]{architecture-mda-lod-ext-general.pdf}
  \end{center}
  \begin{center}
      \textbf{Abbreviations}\\[1ex]\scriptsize
      T-Module is Transformation module,
      MDA is Model-Driven Architecture,
      CIM is Computationally Independent Model,
      PIM is Platform Independent Model,
      PSM is Platform Specific Model,
      T-Box is Terminological Box,
      A-Box is Instance Box,
      A-Box is Instance Box,
      NGS is Next-Generation Sequencing,
      DB is Database.
  \end{center}

@@

@@

\section{Front matter}

@@

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Semantic Web technologies and standards are becoming one of productive assets for software development and integration on various levels of design and implementation.  These include OSI level 6 data representation, database and file storage formats, the source data for indexes, hypertext markup, source data for document authoring, representing abstract software models, \emph{etc}.  They also smoothly combined with development and data analysis tools, giving raise a new common-ground environment for realizing integration on the semantic level.

@@

KGs changed the view point to the knowledge base as being a part of whole totality of knowledge, implying the obeying the global standards and techniques of its acquisition and processing \cite{hogan}.  The following outlines their definitions \cite{hogan}.
 \begin{itemize}
  \item Notions \emph{data} and \emph{knowledge} are converged into ``something is \emph{known}'';
  \item KG contains, in general, data, relations, and metadata;
  \item The same KG is being filled in and processed by different independent modules, \emph{e.g.}, withing SPARQL queries with UPDATE statements;
  \item KG fundamental software development allow to \emph{postpone} the formal definition of a data schema;
  \item There is three points of view to graph schemata: \emph{semantic}, which is aimed at generalization, \emph{validating}, reflecting semantic verification, checking completeness w.r.t. sets of predefined general relations, and \emph{emergent}, observed or deduced from exploitation experience, as a set of new generalized structures and \emph{refactoring} of the KG.
  \end{itemize}
With these general constraints, design of application modules are to be built on standardized data interchange formats, which are to be easily converted from and to KG representation, as well as applications are to be constructed out of service modules supporting SW endpoints (SPARQL) whenever possible.  Our applications are composition of modules showed in Figure~\ref{fig:arch-services}.
\begin{figure}[bth]
  \begin{center}
    \includegraphics[width=0.9\linewidth]{architecture-mda-lod-ext-general.pdf}
  \end{center}
  \begin{center}
      \textbf{Abbreviations}\\[1ex]\scriptsize
      T-Module is Transformation module,
      MDA is Model-Driven Architecture,
      CIM is Computationally Independent Model,
      PIM is Platform Independent Model,
      PSM is Platform Specific Model,
      T-Box is Terminological Box,
      A-Box is Instance Box,
      A-Box is Instance Box,
      NGS is Next-Generation Sequencing,
      DB is Database.
  \end{center}

@@

The kernel module of applications is \texttt{KG server} (A2) built on the base of ClioPatria \cite{cliopatria} with web-server and Pengines plug-ins.  ClioPatria is a web-application implemented in SWI-Prolog \cite{swi} being ISO-compliant Prolog implementation.  KG functions are realized with SW package, supplied in standard SWI-Prolog distribution.  Pengines service enables SW web-applications requesting logical inference on the server side \cite{pengines} with integration of browser JavaScript environment for supporting input-output.  Pengines allows integration of Prolog programs as its plugins.  Kernel module stores A-~and T-boxes of applications, \emph{i.e.} it is application database.  Other modules support application (C3) with frequently used services.

Authoring tool (B1) is used to provide document generation and integration \cite{sibiricon}.  This tool enables document content and structure inference in browser utilizing templates and data from other web pages and websites.  Programmer implements a view of an object as a web page with an extended version of RDFa.  The page are generated with any programming system, \emph{e.g.} PHP.  Authoring tool engine collects web page chunks and composing the document out of these chunks as a XHTML page.  This composed page can be stored in KG as set of triples and indexed in Text indexing module (C1), printed with browser's printing engine, or converted to PDF.  The tool allow user to change document content in WISYWIG mode in some extent.  Its general structure is defined by changing template at server side.

Text indexing engine (C1) provides service for storing documents represented as KG triples, JSONs, and BLOBS of any format with corresponding text layers.  This module has two implementations.  One is build on Sphinx Search engine, and the second one is Elasticsearch.  Elasticsearch supports JSON as the only document format and easily integrated with KG using JSON-LD for triple representation.  Sphinx Search is much faster in text and database records indexing and consumes less RAM as implemented in C++.  The choice of an engine depends on the primary document representation to be used in application.

BLOB stored documents having no markup usually contain useful data, which are to be revealed for application use.  They could be report data or scientific paper content represented as PDF, DJVU files or raster images. Data recognition is realized in PDF/Text document processing module (C2).  Data processing workers query Text indexing module for BLOB documents and add recognized text and high-order structure layers.  All obtained layers are stored in indexing database of C1, as well as data of interest can be converted in triples and sent to KG.

Processing KG for the various purposed are presented as KB processing module (B3).

@@

@@

\bibitem{cliopatria}
\bibitem{sibircon} Logical objects

### 2021-12-05 talk-aicts-2021-12-06.tex
Added new page to paper.
Presentation should be ok now

\title{Technologies of Semantic WEB as an environment of application development and integration}
\author[E.~Cherkashin]{\bfseries%
  Evgeny Cherkashin}
\institute{\normalsize Matrosov Institute for System Dynamics and Control Theory of Siberian Branch of Russian Academy of Sciences, Irkutsk, Russia\\%
  \email{\href{mailto:eugeneai@icc.ru}{eugeneai@icc.ru}}%
}
\date[2021]{AICTS'2021, December, 06, 2021 \\
Irkutsk, Russia}
%\date{\today}
\maketitle

@@

The following aspects are under consideration:
  \begin{enumerate}
  \item Ontological model representation
  \item Application data representation
  \item Model transformation
  \item Document publication
  \item Application integration via knowledge graph data documents
  \end{enumerate}
\end{frame}

@@

\begin{frame}[fragile]
  \frametitle{Data formats for graph representation}
  \begin{columns}
    \begin{column}{0.5\linewidth}
\begin{itemize}
\item N-Triples
\begin{minted}[fontsize=\tiny]{text}
<http://mythology.Greek.org/#Cronus>
    <http://www.example.org/schemas/relationship/fatherOf>
    <http://mythology.Greek.org/#Zeus>.
\end{minted}
\item Turtle
\begin{minted}[fontsize=\tiny]{text}
@prefix rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#> .
@prefix dc: <http://purl.org/dc/elements/1.1/> .
@prefix ex: <http://example.org/stuff/1.0/> .
<http://www.w3.org/TR/rdf-syntax-grammar>
  dc:title "RDF/XML Syntax Specification (Revised)" ;
  ex:editor [
    ex:fullname "Dave Beckett";
    ex:homePage <http://purl.org/net/dajobe/>
  ] .
\end{minted}
\item Notation 3 (N3)
\begin{minted}[fontsize=\tiny]{text}
@prefix dc: <http://purl.org/dc/elements/1.1/> .
<http://en.wikipedia.org/wiki/Tony_Benn>
  dc:title "Tony Benn" ;
  dc:publisher "Wikipedia" .
\end{minted}
\end{itemize}
\end{column}
\begin{column}{0.5\linewidth}
\begin{itemize}
\item RDF/XML
\begin{minted}[fontsize=\tiny]{xml}
<rdf:RDF
    xmlns:rdf="http://www.w3.org/1999/02/22-rdf-syntax-ns#"
    xmlns:dc="http://purl.org/dc/elements/1.1/">
  <rdf:Description rdf:about="http://en.wikipedia.org/wiki/Tony_Benn">
    <dc:title>Tony Benn</dc:title>
    <dc:publisher>Wikipedia</dc:publisher>
  </rdf:Description>
</rdf:RDF>
\end{minted}
\item JSON-LD
\begin{minted}[fontsize=\tiny]{jsonld}
{
  "@context": {
    "name": "http://xmlns.com/foaf/0.1/name",
    "homepage": {
      "@id": "http://xmlns.com/foaf/0.1/workplaceHomepage",
      "@type": "@id"
    },
    "Person": "http://xmlns.com/foaf/0.1/Person"
  },
  "@id": "http://me.markus-lanthaler.com",
  "@type": "Person",
  "name": "Markus Lanthaler",
  "homepage": "http://www.tugraz.at/"
}
\end{minted}
\end{itemize}
\end{column}
\end{columns}
\end{frame}

@@

\begin{frame}[fragile]
  \frametitle{Ontological instruments:  ClioPatria}
  \begin{center}
    \includegraphics[width=0.8\linewidth]{pics/ClioPatria-shot.png}
  \end{center}
\end{frame}

@@

%\begin{frame}
%  \frametitle{Knowledge Graphs}
For the Knowledge Graphs (KG), the following is of interest.
 \begin{itemize}
  \item Converged notions \textbf{data} and \textbf{knowledge} as something is \textbf{known}
  \item Contain data, relations, and metadata (vocabularies)
  \item Distinguished \textbf{node filling in} and \textbf{processing} graph triples, \emph{e.g.}, withing SPARQL queries with UPDATEs
  \item Allow \textbf{postpone} the formal definition of a schema
  \item Three types of graph schemata: \textbf{semantic} (aimed at generalization), \textbf{validating} (\textbf{e.g.} semantics, \textbf{completeness} w.r.t. sets of relations), and \textbf{emergent} (infer a set of generalized structures and \textbf{reconstruct} the KG).
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Information infrastructure for supporting Baikal microbiome research}
  \centering
  \includegraphics[width=\linewidth]{microbiome.png}
\end{frame}

\begin{frame}
  \frametitle{Microbiome study aims}
  \centering
  \includegraphics[width=\linewidth]{microbiome-study.png}
\end{frame}

\begin{frame}
  \frametitle{Microbiome study process}
  \centering
  \includegraphics[width=\linewidth]{microbiome-study2.png}
\end{frame}

@@

\begin{frame}
  \frametitle{The process of data analysis (MiSeq SOP)}
  \begin{enumerate}
  \item Reconstruct \emph{cotigs} (contiguous gene parts) from ``left'' and ``right'' \emph{readings}.
  \item Trim \textbf{bar-code} and other \emph{primers}.
  \item Filter sequences according to formal criteria (ambiguity, average length, maximal length of homopolymer).
  \item Classify \emph{unique} sequences and count their appearance in groups (samples).
  \item Alignment with reference sequences from SILVA database.
  \item Filter non-hanging sequences.
  \item Filter chimeras, fund unique sequences again.
  \item Classify sequences with respect to existing taxa hierarchy. Get \textbf{OTU}s.
  \end{enumerate}

  After these stages a large number of OTU\footnote{Operation Taxonomic Unit} classified has been obtained.
\end{frame}

@@

class MothurChimeraCcodeOperator extends MothurGeneratedOperator {
  private InputPort fastaInPort = getInputPorts().createPort("fasta");
  private InputPort referenceInPort = getInputPorts().createPort("reference");
  private OutputPort chimeraOutPort = getOutputPorts().createPort("chimera");
  private OutputPort mapinfoOutPort = getOutputPorts().createPort("mapinfo");
  private OutputPort accnosOutPort = getOutputPorts().createPort("accnos");

@@

[] a gal:Suite ;
 ngsp:module [ a gal:Module,
     ngsp:Module ;
   gal:command " ## . . . .  ";
   gal:exit_code [ gal:level "fatal" ;
     gal:range "1:" ] ;
       gal:inputs [ gal:checked "false" ;
         gal:conditional [ gal:param [ gal:help "" ;
           gal:option [ gal:value "yes" ;
              dc:description "yes" ],
                [ gal:value "no" ;
                  dc:description "no" ] ;
                dc:description "Trim with an oligos file?" ;
                dc:title "add" ;
              rdfs:range "select" ] ;
           gal:when [ gal:value "no" ],
           [ gal:param [ gal:min "0" ;
             gal:value "0" ;
             dc:description "pdiffs - number of differences to allow in the primer (default 0)" ;
             dc:title "pdiffs" ;
             rdfs:range "integer" ],
             [ gal:min "0" ;
               gal:value "0" ;
               dc:description "bdiffs - number of differences to allow in the barcode (default 0)" ;
               dc:title "bdiffs" ;
               rdfs:range "integer" ],
             [ gal:min "0" ;
               gal:value "0" ;
               dc:description "tdiffs - total number of differences to allow in primer and barcode (default 0)" ;
             dc:title "tdiffs" ;
             rdfs:range "integer" ] ] ] ]
   dc:identifier "mothur_make_contigs" ;
   dc:title "Make.contigs",
            "make.contigs" ;
   schema:sku 1 ''

@@

@@

\begin{frame}[fragile]\frametitle{Architecture of services}
  \begin{columns}
    \begin{column}{0.7\linewidth}
      \includegraphics[width=1\linewidth]{architecture-mda-lod-ext-galaxy.pdf}
    \end{column}
    \begin{column}{0.3\linewidth}
      \textbf{Abbreviations}\\[1ex]\scriptsize
      T-Module is Transformation module\\
      MDA is Model-Driven Architecture\\
      CIM is Computationally Independent Model\\
      PIM is Platform Independent Model\\
      PSM is Platform Specific Model\\
      T-Box is Terminological Box\\
      A-Box is Instance Box\\
      A-Box is Instance Box\\
      NGS is Next-Generation Sequencing\\
      DB is Database
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Used ontologies}

  Standardized ontologies

  \begin{itemize}
  \item Friend-of-a-friend (\textbf{foaf}) for agent information: individuals, legal entities, program agents.
  \item Provenance (\textbf{prov}) for making references between documents.
  \item Dublin Core (\textbf{dc}) for published resource metadata mark up.
  \item DBPedia resource (\textbf{dbr}) to refer external classes and instance objects.
  \item Schema.org (\textbf{schema}) for Google, Yandex, Yahoo, \emph{etc}. searchable objects, structural elements.
  \item The Bibliographic Ontology (\textbf{bibo}) used for literature reference mark up.
  \item Open annotation (\textbf{oa}) as an ``bookmark'' ontology.
  \end{itemize}

  Non-standard ontologies

@@

\begin{frame}
  \frametitle{Future activities}
  The future activities supposed to be as follows:
  \begin{enumerate}
  \item Having dataflow models of MiSeq SOP and other techniques, device an intelligent subsystem, which will construct computational procedures for a predefined set of data processing tasks (\emph{AI's problem solving}).
  \item Integrate dataflow and data storage with Galaxy.
    \begin{itemize}
       \item Realize adapters of data/metadata storage and retrieving, as well as the storage.
       \item Create a more sophisticated source code parser or PIM model of computation so we will able to infer metadata for the output on the metadata of input (partially done in November).
         % \item Adapt our ``pattern-directed'' approach of semantically marked up document authoring to present input and output to the scientific communities.
       \item Adapt our document authoring tools to Galaxy allowing LOD representation of results.
  \end{itemize}
  \item Implement integration to biological/gene databases.
  \item Write a handbook on Logtalk programming strategies with its author Paulo Mora.
  \end{enumerate}
\end{frame}

@@

\begin{frame}
  \frametitle{Document authoring and storage}
  In most cases documents are created as a result of
  \begin{itemize}
  \item creative activity of a person with a text processors (authoring);
  \item printing a digital copy or a data record in a database;
  \item aggregation operation over database records (report).
  \end{itemize}
  Then it is stored either as a physical paper and/or a digital document (PDF, DOCX, HTML).

@@

\begin{frame}
\frametitle{Cartographical WEB-service with knowledge graph of South-Siberian faults}
\textbf{Aim} is to construct a WEB-GIS browser for faults stored in the KG.
\begin{itemize}
\item Scalability to external data with converters (TODO)
\item Interdisciplinary data representation
\item Application development with nowadays WEB techniques
\item Digital platform for data publication in ``Digital Baikal'' project
\end{itemize}
% \end{frame}

@@

% <<Загрузка>> онтологий
schema(Prop, 'of', SubjName):-     % used on translation stage
        var(SubjName),
        geob_prop(Prop,_).

@@

\begin{frame}
  \frametitle{Conclusion}
  The following results have been obtained as for today:
  \begin{itemize}
  \item A technique for \textbf{domain} model representation has been developed and tested.
  \item A programming technique using object-oriented logical language Logtalk is devised.
  \item Prototypes of various transformation procedures are implemented.
  \item Transformation tools are tested in application areas and no significant technical problems were mentioned.
  \end{itemize}
  Further development directions are as follows:
  \begin{itemize}
  \item A technique for document automatic markup with vocabulary entities.
  \item A transformation implemenation techniques, minimizing usage of dynamic objects, targeting on macro properties of Logtalk.
  \item Form a toolset out of existing prototypes obeying nowadays software development requirements.
  \end{itemize}
  The source codes are available at \url{https://github.com/isu-enterprise/icc.xmitransform}, \url{https://github.com/eugeneai/icc.mothurpim}.
\end{frame}

\begin{frame}
  \frametitle{Technologies used (open source)}
  Python-3.x.x (\url{http://python.org}),\\
  ZCA (\url{https://muthukadan.net/docs/zca.html}),\\
  SWIG (\url{http://swig.org/}),\\
  SWI-Prolog (\url{https://www.swi-prolog.org/}),\\
  Logtalk (\url{https://logtalk.org/}),\\
  ClioPatria (\url{https://cliopatria.swi-prolog.org/home}),\\
  Pengines (\url{https://pengines.swi-prolog.org/docs/index.html}),\\
  LOV (\url{https://lov.linkeddata.es/dataset/lov/}),\\
  Elastic Search (\url{https://www.elastic.co/}),\\
  Kyotocabinet (\url{https://fallabs.com/kyotocabinet/}),\\
  DBPedia (\url{https://wiki.dbpedia.org/}),\\
  Mothur (\url{https://mothur.org/}),\\
  Galaxy (\url{https://usegalaxy.org/}),\\
  R (\url{https://www.r-project.org/}), \\
  Dust.js (\url{https://akdubya.github.io/dustjs/})
\end{frame}

>>>>

\title{Technologies of Semantic WEB as an environment of application development and integration}
\author[E.~Cherkashin]{\bfseries%
  Evgeny Cherkashin, Nikita Dorodnykh, Alexey Shigarov, \emph{et al.}}
\institute{\normalsize Matrosov Institute for System Dynamics and Control Theory of Siberian Branch of Russian Academy of Sciences, Irkutsk, Russia\\%
  \email{\href{mailto:eugeneai@icc.ru}{eugeneai@icc.ru}}%
}
\date[2021]{AICTS'2021, December, 06, 2021 \\
Irkutsk, Russia}
%\date{\today}
\maketitle

@@

The following aspects are under consideration:
  \begin{enumerate}
  \item Application data representation
  \item Ontological model representation
  \item Document publication
  \item Application integration % via knowledge graph data documents
  \item Model transformation
  \end{enumerate}
\end{frame}

@@

\begin{frame}[fragile]
  \frametitle{Data formats for graph representation}
  \begin{columns}
    \begin{column}{0.5\linewidth}
\begin{itemize}
\item N-Triples
\begin{minted}[fontsize=\tiny]{text}
<http://mythology.Greek.org/#Cronus>
    <http://www.example.org/schemas/relationship/fatherOf>
    <http://mythology.Greek.org/#Zeus>.
\end{minted}
\item Turtle
\begin{minted}[fontsize=\tiny]{text}
@prefix rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#> .
@prefix dc: <http://purl.org/dc/elements/1.1/> .
@prefix ex: <http://example.org/stuff/1.0/> .
<http://www.w3.org/TR/rdf-syntax-grammar>
  dc:title "RDF/XML Syntax Specification (Revised)" ;
  ex:editor [
    ex:fullname "Dave Beckett";
    ex:homePage <http://purl.org/net/dajobe/>
  ] .
\end{minted}
\item Notation 3 (N3)
\begin{minted}[fontsize=\tiny]{text}
@prefix dc: <http://purl.org/dc/elements/1.1/> .
<http://en.wikipedia.org/wiki/Tony_Benn>
  dc:title "Tony Benn" ;
  dc:publisher "Wikipedia" .
\end{minted}
\end{itemize}
\end{column}
\begin{column}{0.5\linewidth}
\begin{itemize}
\item RDF/XML
\begin{minted}[fontsize=\tiny]{xml}
<rdf:RDF
    xmlns:rdf="http://www.w3.org/1999/02/
               22-rdf-syntax-ns#"
    xmlns:dc="http://purl.org/dc/elements/1.1/">
  <rdf:Description rdf:about="http://en.wikipedia.org/
                              wiki/Tony_Benn">
    <dc:title>Tony Benn</dc:title>
    <dc:publisher>Wikipedia</dc:publisher>
  </rdf:Description>
</rdf:RDF>
\end{minted}
\item JSON-LD
\begin{minted}[fontsize=\tiny]{jsonld}
{
  "@context": {
    "name": "http://xmlns.com/foaf/0.1/name",
    "homepage": {
      "@id": "http://xmlns.com/foaf/0.1/
              workplaceHomepage",
      "@type": "@id"
    },
    "Person": "http://xmlns.com/foaf/0.1/Person"
  },
  "@id": "http://me.markus-lanthaler.com",
  "@type": "Person",
  "name": "Markus Lanthaler",
  "homepage": "http://www.tugraz.at/"
}
\end{minted}
\end{itemize}
\end{column}
\end{columns}
\end{frame}

@@

\begin{frame}[fragile]
  \frametitle{Ontological instruments:  ClioPatria}
  \begin{center}
    \includegraphics[width=0.8\linewidth]{pics/ClioPatria-shot.png}
  \end{center}
  Using standard vocabularies form cross-application platform, \emph{e.g.}, interpreting relations.
\end{frame}

@@

%\begin{frame}
%  \frametitle{Knowledge Graphs}
For the Knowledge Graphs (KG), the following is of interest.
 \begin{itemize}
  \item Converged notions \textbf{data} and \textbf{knowledge} as something is \textbf{known}
  \item Contain data, relations, and metadata (vocabularies)
  \item Distinguished \textbf{node filling in} and \textbf{processing} graph triples, \emph{e.g.}, with SPARQL queries with UPDATEs
  \item Allow \textbf{postpone} the formal definition of a schema
  \item Three types of graph schemata: \textbf{semantic} (aimed at generalization), \textbf{validating} (\textbf{e.g.} semantics, \textbf{completeness} w.r.t. sets of relations), and \textbf{emergent} (infer a set of generalized structures and \textbf{reconstruct} the KG).
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Useful standard vocabularies}

  Standardized vocabularies

  \begin{itemize}
  \item Friend-of-a-friend (\textbf{foaf}) for agent information: individuals, legal entities, program agents.
  \item Provenance (\textbf{prov}) for making references between documents.
  \item Dublin Core (\textbf{dc}) for published resource metadata mark up.
  \item DBPedia resource (\textbf{dbr}) to refer external classes and instance objects.
  \item Open annotation (\textbf{oa}) as an ``bookmark'' ontology.
  \item The Bibliographic Ontology (\textbf{bibo}) used for literature reference mark up.
  \item Schema.org (\textbf{schema}) for Google, Yandex, Yahoo, \emph{etc}. searchable objects, structural elements.
  \end{itemize}

  Non-standard vocabularies

\begin{frame}
  \frametitle{Applications}
  \vfil
  \begin{center}
    \Large Applications
  \end{center}
  \vfil
\end{frame}

\begin{frame}
  \frametitle{Application: Information infrastructure for supporting Baikal microbiome research}
  \centering
  \includegraphics[width=\linewidth]{microbiome.png}
\end{frame}

% \begin{frame}
%   \frametitle{Microbiome study aims}
%   \centering
%   \includegraphics[width=\linewidth]{microbiome-study.png}
% \end{frame}

% \begin{frame}
%   \frametitle{Microbiome study process}
%   \centering
%   \includegraphics[width=\linewidth]{microbiome-study2.png}
% \end{frame}

@@

% \begin{frame}
%   \frametitle{The process of data analysis (MiSeq SOP)}
%   \begin{enumerate}
%   \item Reconstruct \emph{cotigs} (contiguous gene parts) from ``left'' and ``right'' \emph{readings}.
%   \item Trim \textbf{bar-code} and other \emph{primers}.
%   \item Filter sequences according to formal criteria (ambiguity, average length, maximal length of homopolymer).
%   \item Classify \emph{unique} sequences and count their appearance in groups (samples).
%   \item Alignment with reference sequences from SILVA database.
%   \item Filter non-hanging sequences.
%   \item Filter chimeras, fund unique sequences again.
%   \item Classify sequences with respect to existing taxa hierarchy. Get \textbf{OTU}s.
%   \end{enumerate}

%   After these stages a large number of OTU\footnote{Operation Taxonomic Unit} classified has been obtained.
% \end{frame}

@@

class MothurChimeraCcodeOperator extends MothurGeneratedOperator {
  private InputPort fastaInPort = getInputPorts().createPort("fasta");
  private InputPort referenceInPort = getInputPorts().
                                         createPort("reference");
  private OutputPort chimeraOutPort = getOutputPorts().createPort("chimera");
  private OutputPort mapinfoOutPort = getOutputPorts().createPort("mapinfo");
  private OutputPort accnosOutPort = getOutputPorts().createPort("accnos");

@@

[] a gal:Suite ;
 ngsp:module [ a gal:Module,
     ngsp:Module ;
   gal:command " ## . . . .  ";
   gal:exit_code [ gal:level "fatal" ;
     gal:range "1:" ] ;
       gal:inputs [ gal:checked "false" ;
         gal:conditional [ gal:param [ gal:help "" ;
           gal:option [ gal:value "yes" ;
              dc:description "yes" ],
                [ gal:value "no" ;
                  dc:description "no" ] ;
                dc:description "Trim with an oligos file?" ;
                dc:title "add" ;
              rdfs:range "select" ] ;
           gal:when [ gal:value "no" ],
           [ gal:param [ gal:min "0" ;
             gal:value "0" ;
             dc:description "pdiffs - number of differences . . . . . ." ;
             dc:title "pdiffs" ;
             rdfs:range "integer" ],
             [ gal:min "0" ;
               gal:value "0" ;
               dc:description "bdiffs - number of differences . . . . . . " ;
               dc:title "bdiffs" ;
               rdfs:range "integer" ],
             [ gal:min "0" ;
               gal:value "0" ;
               dc:description "tdiffs - total number of diffe... . . . . ." ;
             dc:title "tdiffs" ;
             rdfs:range "integer" ] ] ] ]
   dc:identifier "mothur_make_contigs" ;
   dc:title "Make.contigs",
            "make.contigs" ;
   schema:sku 1 ''

@@

\begin{frame}[fragile]\frametitle{Architecture of services for NGS}
  \begin{columns}
    \begin{column}{0.7\linewidth}
      \includegraphics[width=1\linewidth]{architecture-mda-lod-ext-general.pdf}
    \end{column}
    \begin{column}{0.3\linewidth}
      \textbf{Abbreviations}\\[1ex]\scriptsize
      T-Module is Transformation module\\
      MDA is Model-Driven Architecture\\
      CIM is Computationally Independent Model\\
      PIM is Platform Independent Model\\
      PSM is Platform Specific Model\\
      T-Box is Terminological Box\\
      A-Box is Instance Box\\
      A-Box is Instance Box\\
      NGS is Next-Generation Sequencing\\
      DB is Database
    \end{column}
  \end{columns}
\end{frame}

@@

@@

% \begin{frame}
%   \frametitle{Future activities}
%   The future activities supposed to be as follows:
%   \begin{enumerate}
%   \item Having dataflow models of MiSeq SOP and other techniques, device an intelligent subsystem, which will construct computational procedures for a predefined set of data processing tasks (\emph{AI's problem solving}).
%   \item Integrate dataflow and data storage with Galaxy.
%     \begin{itemize}
%        \item Realize adapters of data/metadata storage and retrieving, as well as the storage.
%        \item Create a more sophisticated source code parser or PIM model of computation so we will able to infer metadata for the output on the metadata of input (partially done in November).
%          % \item Adapt our ``pattern-directed'' approach of semantically marked up document authoring to present input and output to the scientific communities.
%        \item Adapt our document authoring tools to Galaxy allowing LOD representation of results.
%   \end{itemize}
%   \item Implement integration to biological/gene databases.
%   \item Write a handbook on Logtalk programming strategies with its author Paulo Mora.
%   \end{enumerate}
% \end{frame}

@@

\begin{frame}
  \frametitle{Application: Document authoring and storage}
  In most cases documents are created as a result of
  \begin{itemize}
  \item creative activity of a person with a text processors (authoring);
  \item printing a digital copy or a data record in a database;
  \item aggregation operation over database records (report).
  \end{itemize}
  Then it is stored either as a physical paper and/or a digital document (PDF, DOCX, HTML).

@@

\begin{frame}
\frametitle{Application: Cartographical WEB-service with knowledge graph of South-Siberian faults}
\textbf{Aim} is to construct a WEB-GIS browser for faults stored in the KG.
\begin{itemize}
\item Scalability to external data with converters (TODO)
\item Interdisciplinary data representation
\item Application development with nowadays WEB techniques
\item Digital platform for data publication in ``Digital Baikal'' project
\end{itemize}
% \end{frame}

@@

% Interpret any well described relation
% between a subject (fault) via ''of''.
schema(Prop, 'of', SubjName):-     % used on translation stage
        var(SubjName),
        geob_prop(Prop,_).

@@

% \begin{frame}
%   \frametitle{Conclusion}  % On Web GIS
%   The following results have been obtained as for today:
%   \begin{itemize}
%   \item A technique for \textbf{domain} model representation has been developed and tested.
%   \item A programming technique using object-oriented logical language Logtalk is devised.
%   \item Prototypes of various transformation procedures are implemented.
%   \item Transformation tools are tested in application areas and no significant technical problems were mentioned.
%   \end{itemize}
%   Further development directions are as follows:
%   \begin{itemize}
%   \item A technique for document automatic markup with vocabulary entities.
%   \item A transformation implemenation techniques, minimizing usage of dynamic objects, targeting on macro properties of Logtalk.
%   \item Form a toolset out of existing prototypes obeying nowadays software development requirements.
%   \end{itemize}
%   Projects sources are at \url{https://github.com/isu-enterprise} and \url{https://github.com/eugeneai}.
% \end{frame}

\begin{frame}
  \frametitle{Conclusion}  % On Web GIS
  Web-3.0 (SW and KG) are convenient and productive basis of application development, allowing one to
  \begin{itemize}
  \item Integrate on various levels of software representation, \emph{i.e.}, from data to abstract models.
  \item There are assets to utilize global resources in applications (\texttt{DBPedia.org}).
  \item The inherited software are involved via adapters.
  \end{itemize}
  In our projects we has been developed techniques to
  \begin{itemize}
  \item Data processing and model transformation using Prolog-based logical inference.
  \item Logtalk is a perfect instrument for knowledge encapsulation and processing, complex data synthesis.
  \end{itemize}
\end{frame}

% \begin{frame}
%   \frametitle{Technologies used (open source)}
%   Python-3.x.x (\url{http://python.org}),\\
%   ZCA (\url{https://muthukadan.net/docs/zca.html}),\\
%   SWIG (\url{http://swig.org/}),\\
%   SWI-Prolog (\url{https://www.swi-prolog.org/}),\\
%   Logtalk (\url{https://logtalk.org/}),\\
%   ClioPatria (\url{https://cliopatria.swi-prolog.org/home}),\\
%   Pengines (\url{https://pengines.swi-prolog.org/docs/index.html}),\\
%   LOV (\url{https://lov.linkeddata.es/dataset/lov/}),\\
%   Elastic Search (\url{https://www.elastic.co/}),\\
%   Kyotocabinet (\url{https://fallabs.com/kyotocabinet/}),\\
%   DBPedia (\url{https://wiki.dbpedia.org/}),\\
%   Mothur (\url{https://mothur.org/}),\\
%   Galaxy (\url{https://usegalaxy.org/}),\\
%   R (\url{https://www.r-project.org/}), \\
%   Dust.js (\url{https://akdubya.github.io/dustjs/})
% \end{frame}
